CUNY SPSS Data Science Math Homework 2

### 2018-07-22 data-science-math/Homework 2.Rmd
Fixed table in question 24
Finished Homework 2

$$P(l) =\frac{\{1,2,3,4,5\}}{\{1,2,3,4,5,6\}}=5/6\approx0.8333$$
```{r}
5/6
```

@@

@@

q11_data <- tidyr::crossing(drinks, sandwiches) %>%
  tidyr::crossing(., chips)
nrow(q11_data)
```

@@

$_9 P_4$

@@

$_11 C_8$

@@

$\frac{_{12} P_8}{_{12} C_4}$

@@

----------------------------------------
  &nbsp;    NA    NA    NA    NA    NA  
---------- ----- ----- ----- ----- -----
  **x**      5     6     7     8     9

@@

$$\sigma^2=\frac{\Sigma (x-E(x))^2}{N} = \frac{(5-7.2)^2+(6-7.2)^2+(7-7.2)^2+(8-7.2)^2+(9-7.2)^2}{5}$$
$$\sigma^2= \frac{4.84+1.44+0.04+0.64+3.24}{5}=\frac{10.2}{5}=2.04$$

```{r}
v <- sum((x-ex)**2)/length(x)
v
```
#### Step 3

Find the standard deviation. Round your answer to one decimal place.
$$\sigma = \sqrt{\sigma ^2} = \sqrt{2.04} \approx 1.4$$
```{r}
v**(1/2)
```

@@

Find the value of $P(X\hat e9)$. Round your answer to one decimal place.

@@

Find the value of $P(X\c{C} 7)$. Round your answer to one decimal place.

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

$$P(l) =\frac{\{1,2,3,4,5\}}{\{1,2,3,4,5,6\}}=\frac{5}{6}\approx0.8333$$
```{r}
5/6
```

@@

&nbsp;

@@

q11_data <- crossing(drinks, sandwiches) %>%
  crossing(., chips)
nrow(q11_data)
```

@@

$$_9 P_4$$
$$_9 P_4=\frac{9!}{(9-4)!}=\frac{362880}{120}=3024$$
```{r}
N <- 9
n <- 4
factorial(N)/factorial((N-n))
```

@@

$$_{11}C_8$$

$$_{11}C_8=\frac{11!}{8!\cdot(11-8)!}=\frac{11!}{8!\cdot3!}=\frac{39916800}{403208 \cdot 6}=165$$

```{r}
N <- 11
n <- 8
factorial(N)/(factorial(n)*factorial(N-n))
```

@@

$$\frac{_{12} P_8}{_{12} C_4}$$

$$_{12}P_8 = \frac{12!}{(12-8)!}=\frac{479001600}{4!}=\frac{479001600}{24}=19958400$$

$$ _{12}C_4=\frac{12!}{4!\cdot(12-4)!}=\frac{479001600}{24\cdot8!}=\frac{479001600}{24\cdot 40320}=\frac{479001600}{967680}=495$$
$$\frac{_{12} P_8}{_{12} C_4}=\frac{19958400}{495}=40320$$

```{r}
N <- 12
n1 <- 8
n2 <- 4
q21_p <- factorial(N)/factorial((N-n1))
q21_c <- factorial(N)/(factorial(n2)*factorial(N-n2))
q21_p/q21_c
```

@@

----------------------------------------

---------- ----- ----- ----- ----- -----
  **x**      5     6     7     8     9

@@

$$\sigma^2=\Sigma_i (x_i-E(x))^2P(x_i) $$
$$\sigma^2= 0.1(5-7.2)^2+0.2(6-7.2)^2+0.3(7-7.2)^2+0.2(8-7.2)^2+0.2(9-7.2)^2$$
$$\sigma^2= (0.1\times 4.84)+(0.2 \times 1.44)+(0.3 \times 0.04)+(0.2 \times 0.64)+(0.2 \times 3.24)$$
$$\sigma^2=0.484+0.288+0.012+0.128+0.648\approx1.6$$

```{r}
v <- sum(px*(x-ex)**2)
v
```

&nbsp;

#### Step 3

Find the standard deviation. Round your answer to one decimal place.
$$\sigma = \sqrt{\sigma ^2} \approx \sqrt{1.6} \approx 1.3$$
```{r}
v**(1/2)
```
The difference is due the the fact that I used 1.6 as the variance instead of 1.56.

&nbsp;

@@

Find the value of $P(X\geq9)$. Round your answer to one decimal place.

The answer is **0.2** because this is the only P(x) where x >= 9.

```{r}
sum(px[x>=9])
```

&nbsp;

@@

Find the value of $P(X\leq7)$. Round your answer to one decimal place.

$$0.1 + 0.2 + 0.3= 0.6$$

```{r}
sum(px[x<=7])
```

@@

The player shooting $\frac{188}{376}$ free throws so they have a 50% probability of making the shot.  We can now estimate the probability of them making the next 3 shots as $\frac{1}{8}$.  The probability that they won't is $\frac{7}{8}$.

The expected value of the proposition is:

$$P(\mbox{winning bet}) \times \mbox{winnings } - P(\mbox{not winning bet}) \times \mbox{loss}$$

$$ \frac{1}{8}\times 23 - \frac{7}{8} \times 4 = \frac{23}{8}-\frac{28}{8} = -\frac{5}{8} \approx -0.63$$
```{r}
makes_it <-c(1,0)

q25_data <- crossing(makes_it, makes_it) %>%
  crossing(., makes_it)

q25_data$shots_made <- rowSums(q25_data)

made_all_3 <- 1
n <- nrow(q25_data)

pwin <- made_all_3/n
ploose <- 1 - pwin

ex25 <- (pwin * 23) - (ploose * 4)
ex25
```

&nbsp;

@@

$$994 \times -0.63 = -626.22$$

```{r}
994 * ex25
```

@@

```{r}
flipped_tails <- c(1,0)

q26_data <- crossing(flipped_tails, flipped_tails) %>%
  crossing(., flipped_tails) %>%
  crossing(., flipped_tails) %>%
  crossing(., flipped_tails) %>%
  crossing(., flipped_tails) %>%
  crossing(., flipped_tails) %>%
  crossing(., flipped_tails) %>%
  crossing(., flipped_tails) %>%
  crossing(., flipped_tails) %>%
  crossing(., flipped_tails)

q26_data$total_tails <- rowSums(q26_data)

tails_8_or_less <- q26_data %>%
  filter(total_tails < 9) %>%
  nrow(.)
n <- nrow(q26_data)

pwin <- tails_8_or_less/n
ploose <- 1 - pwin

ex26 <- (pwin * 1 )-(ploose * 7)
ex26
```

&nbsp;

@@

```{r}
615 * ex26
```

@@

$$P(c|c) = \frac{12}{51} \cdot \frac{13}{52} \approx 0.059$$

$$ (0.059 \times 583)-((1-0.59)\times 35) \approx 1.46$$

```{r}
pwin <- (12/51)*(13/52)
ploose <- 1 - pwin
ex27 <- (pwin * 583)-(ploose * 35)
ex27
```

The difference is due to the rounding of the probability of wining, but I will use this figure for step 2

&nbsp;

@@

$$632 \times 1.35 = 853.20$$
```{r}
632 * ex27
```

Again the difference is due to rounding.  **Lesson: Don't use rounded values when calculating values (garbage in garbage out).**

@@

$$F(x;p,n) = \sum_{i=0}^{x}{\left( \begin{array}{c} n \\ i \end{array}
 \right) (p)^{i}(1 - p)^{(n-i)}}$$

Not wanting to mess this up I will just us R to solve:

```{r}
p <- 0.3
n <- 10
x <- 2
dbinom(0,n,p)+dbinom(1,n,p)+dbinom(2,n,p)
q28 <- pbinom(x, n, p)
q28
```

The answer is **0.383**.

@@

Given $ p=0.3$ and $n=5$:
$$E = p\cdot n = 1.5$$

```{r}
0.3*5
```

@@

$$F(x;\lambda) = \sum_{i=0}^{x}{\frac{e^{-\lambda}\lambda^{i}} {i!}}$$
Given the complexity of the formula they will only be solved in R.

```{r}
x <- 5
lambda <- 5.5
1 - ppois(1, lambda)
```

The answer is **0.9723**.

@@

```{r}
x <- 4
lambda <- 5.7
1 - ppois(1, lambda)
```

The answer is **0.9776**.

@@

Since they provide a daily lambda the weekly lambda is $7\times 0.4$ or $2.8$.

```{r}
x <- 1
lambda <- 0.4 * 7
ppois(x, lambda)
```

The answer is **0.2311**.

@@

```{r}
q <- 1
m <- 6
n <- 19
k <- 8
1 - phyper(q, m, n, k)
```

The answer is **0.651**.

@@

```{r}
q <- 6
m <- 10
n <- (25-m)
k <- 8
phyper(q, m, n, k)
```

The answer is **0.998**.
